Supporting Information for\\ ``A weakly structured stem for human origins in Africa''

### 2022-08-15 supplement.tex
small changes and formatting

Because CEU and GBR are genetically closely related, we also tested the
consistency of inferred parameters when replacing GBR with Toscani from Italy,
a more distantly related European population set than GBR. Most parameters in
each of the tested models were consistent with parameters inferred using GBR.
The only large difference was found in fitting the continuous migration model,
Nama divergence was $\approx200$ka, rather than $135$ka. 
Our original model using the dataset with GBR
underestimated both $\sigma_{Dz}$ and $\sigma_{\pi_2}$ compared to data
(Figure~\ref{fig:supp-continuous-migration-fits}), and as a result provided a poorer 
fit to the data than the multiple merger model. 
When fitting this same model to the TSI, the model provided a better fit to the $\sigma_{Dz}$ statistics but
a worse fit to the $\sigma_{\pi_2}$ statistics (and still provided a
significantly worse fit to the data compared to the multiple merger model).
Migration rates during
this early period of the model also shifted to larger values between the stem
populations, demonstrating that divergence times and migration rates are
confounded parameters in this model. 
These differences in inferred parameters are likely driven by a combination of factors: the continuous
migration model mis-specifies historical dynamics in the range of 100-200ka;
the TSI are an inappropriate proxy for recent European admixture in East and
South Africa; the TSI may also carry recent North African ancestry, which
is not captured in our model. Finally, there is an amount of confounding between divergence time
and migration rates.

By contrast, the multiple merger model (which
provides a better fit to the data) was inferred to have consistent divergence
and merger times between the models fit using GBR and TSI, with ancestral
populations ``reshuffling'' between 100 and 120ka.

>>>>

Because CEU and GBR are genetically closely related, we also tested the
consistency of inferred parameters when replacing GBR with Toscani from Italy,
a more distantly related European population set than GBR. Most parameters in
each of the tested models were consistent with parameters inferred using GBR.
The only large difference was found in fitting the continuous migration model,
where the Nama divergence was $\approx200$ka rather than $135$ka. Our original
model using the dataset with GBR underestimated both $\sigma_{Dz}$ and
$\sigma_{\pi_2}$ compared to data
(Figure~\ref{fig:supp-continuous-migration-fits}), and as a result provided a
poorer fit to the data than the multiple merger model. When fitting this same
model to the TSI, the model provided a better fit to the $\sigma_{Dz}$
statistics but a worse fit to the $\sigma_{\pi_2}$ statistics (and still
provided a significantly worse fit to the data compared to the multiple
merger model). Migration rates during this early period of the model also
shifted to larger values between the stem populations, demonstrating that
divergence times and migration rates are confounded parameters in this model.
These differences in inferred parameters are likely driven by a combination of
factors: the continuous migration model mis-specifies historical dynamics in
the range of 100-200ka; the TSI are an inappropriate proxy for recent European
admixture in East and South Africa; and the TSI may also carry recent North
African ancestry, which is not captured in our model. By contrast, the
multiple merger model (which provides a better fit to the data) was inferred to
have consistent divergence and merger times between the models fit using GBR
and TSI, with ancestral populations ``reshuffling'' between 100 and 120ka.
